\chapter{The Gospel according to John}

\section{}

In the beginning was reason,
and reason was with the god,
and reason was a god.
It was in the beginning with the god.
All things exist through it,
and nothing exists without it.
In it is life,
and the life is the light of all people.
And the light shines in the darkness,
and the darkness has not overcome it.

There was a man sent from the god
whose name was John. 
He came as a witness to testify to the light,
so that all might believe through him.
He was not that light,
but he had come to testify to the light.
The true light,
which enlightens everyone,
was coming into the cosmos.

It was in the cosmos,
and the cosmos exists through it,
yet the cosmos did not know it.
It came into that which was its home,
and its own people did not accept it.
But to all who did receive it, 
who believed in its name, 
it gave right to be children of the god,
who were born, not of blood 
or of the will of the flesh
or of the will of humanity,
but of the god.

Reason became flesh and lived among us,
and we have seen its glory,
the glory like of that of a father's only son, 
graceful and true.
John testified about it and cried,
`This was the one about whom I said,
``He who comes after me ranks higher than me
because he existed before me.'''
From its fullness we all have received 
grace upon grace.
The law is given through Moses,
and grace and truth is given through Jesus the Messiah.
No man has seen the god at any time.
Only the son, who was at the father's heart,
has known him.

This is the tesimony given by John 
when the Jews sent priests and Levites from Jerusalem to ask him,
`Who are you?'

He confessed and did not deny it,
`I am not the Messiah.'

`What then? Are you Elijah?'

`I am not.'

`Are you the prophet?'

`No.'

`Who are you?  Let us have an answer for those who sent us. 
What do you say about yourself?'

He said, 
`I am the voice of the one crying in the wilderness,
``Straighten the way of the lord.'''
as the prophet Isiah said.

Now those that were sent were from the Pharisees;
they asked him, 
`Why then are you baptising 
if you are neither the Messiah, Elijah, nor the prophet?'

`I baptise with water.  Among you is someone you do not know,
the one who comes after me.  
I am not worthy to untie the strings of his sandal.'

This happened in Bethany, across the Jordan where John was baptising.

The next day, he saw Jesus coming towards him, and he said,
`See, here is the lamb of the god who takes the sin of the world!
This was the one about whom I said,
``He who comes after me ranks higher than me
because he existed before me.''
I did not know him, but I came baptising with water for this very reason,
that he might come to Israel.'
John also testified, saying, 
`I saw the spirit coming down like a dove out of heaven,
it remained on him.
I did not know him, 
but the one who sent me to baptise with water said,
``Whomever you see the spirit come down and stay upon 
is the one who will baptise with the Holy Spirit.''
And I myself have seen and have testified that this is the god's chosen.'

The next day,
John was standing with two of his disciples,
and as he watched Jesus walk by, he cried, 
`Look! Here is the lamb of the god.'
The disciples heard this, and they followed Jesus. 

When Jesus turned and saw them following, he asked,
`What are you looking for?'

They said, `Rabbi, where are you staying?'

`Come, and see.'

They did, and they remained with him that day. 
It was about 16:00.\footnote{The original, `about the tenth hour',
uses Roman time, with starts counting at sunrise (approximately 06:00).
}

One of the two who listened to John speak and followed him was Andrew,
the brother of Simon Peter.
He first found his brother Simon and said,
`We have found the Messiah!'

He brough Simon to Jesus,
who looked at him and said,
`You are Simon son of John; 
you will be called Kephas,
(which is translated Peter.)\footnote{
    Which is further translated to Rock. 
    The given names are from Aramaic \textit{kepha}
    and Greek \textit{petra}, meaning rock.
}

The next day, 
Jesus decided to go to Galilee.
He found Philip and said to him,
`Follow me.'

Philip was from Bethsaida,
the city of Andrew and Peter.
Philip found Nathanael and said, 
`We have found the one whom 
Moses wrote about in the Torah and the Nevi'im: 
Jesus, the son of Joseph, from Nazareth.'

Nathaniel replied, `Can anything good come out of Nazareth?'

`Come, and see.'

When Jesus saw Nathanael coming towards him, he said,
`See, now here is an Israelite with no deceit in him!'

Nathaniel asked, `How do you know me?'

`Before Philip called you, when you were under the fig tree,
I saw you.'

`Rabbi, you are the son of the god and king of Israel!'

`Do you believe because I told you I saw you under the fig tree?
You will see greater things than that.
I am telling you all a solemn truth.
You will see heaven opened, 
and the angels of the god ascending and descending 
to the son of humanity.



%2
\section{}
On the third day, there was a wedding in Cana of Galilee,
and the mother of Jesus was there.
Jesus and his disciples had been invited to the wedding.
When the wine ran out,
the mother of Jesus said,
`They have no more wine.'

Jesus replied, 
`Ma'am, why is that a concern to me and you?
My time has not arrived.'

His mother said to the servants, 
`Do whatever he tells you.'

Standing there were six stone water jars
for ceremonial handwashing,
each holding 75 to 120 litres.\footnote{The Greek says `two or three metretai'.}
Jesus said to them, `Fill the jars with water.'
And they filled them to the brim.

Then he said, `Now, draw some out, and take it to the master of ceremonies.'
They did. 

When the master of ceremonies tasted the water,\footnote{Omitted:
    `\dots that had been turned into wine, not knowing where it came from
    (though the servants who had drawn the water knew)\dots'
} he called the groom and said,
`Usually people serve the good wine first,
and then the worse one
when the guests are drunk.
But you've saved the good wine until now!'

So in Cana of Galilee, 
Jesus did his first sign;
and his disciples believed in him.

After this,
he went down to Capernaum with his mother, his brothers, and his disciples.
They stayed there for a few days.

Passover was near, and Jesus went up to Jerusalem.
Around the temple, he found people selling cattle, sheep, and doves,
and the money changers sitting at their tables.
He made a whip out of cords and forced al of them,
with the sheep and cattle, 
to leave the temple. 
He turned over the 

