\chapter*{Chapter One}
\addcontentsline{toc}{chapter}{Chapter One}

When the god began to make the sky and the earth, 
the earth was a formless nothingness, 
darkness covered the surface of Deep,
and the wind of the god brushed over the surface of the waters. 

Then the god said, 
\enquote{Let there be light.} 
Then there was light. 
The god saw that the light was good. 
The god separated the light from the darkness. 
The god named the light \enquote{day} and the darkness he named \enquote{night}. 
There was an evening, and there was a morning---one day.

Then the god said, 
\enquote{Let there be a cover in the water that will separate water from water.} 
Then the god made the cover and separated the water that was under the cover from the water that was above the cover. 
So it was. 
The god named the cover \enquote{sky}. 
There was an evening, and there was a morning---a second day.

Then the god said, 
\enquote{Let the water under the sky gather into one place, so that dry land may appear.} 
So it was. 
The god named the dry land \enquote{earth} and the gathered waters \enquote{sea}. 
The god saw that it was good. 

Then the god said, 
\enquote{Let the earth bring forth plants: 
ones that make grain as seeds according to their kind and 
trees that make fruits that have in them seeds according to their kind.} 
So it was. 
Thus the earth brought forth plants: 
ones that made grain as seeds according to their kinds and 
trees that made fruits that have in them seeds according to their kind, 
and the god saw that it was good. 
There was an evening, and there was a morning---the third day.

Then the god said, 
\enquote{Let there be lights in the sky, 
to separate day from night, and 
let them be for signs of seasons and days and years, 
and let them be the lights in the sky to illuminate the earth.} 
So it was. 
The god made the two large lights---the bigger one to rule the day 
and the smaller one to rule the night---along with the stars. 
The god put all these in the sky to illuminate the earth, 
to rule over the day and over the night, 
and to separate light from darkness. The god saw that it was good.
There was an evening, and there was a morning---the fourth day.

Then the god said, 
\enquote{Let the waters swarm with swarms of creatures, 
and let the birds fly in the sky above the earth.} 
The god made large sea animals, every living thing that moves in the sea. 
The sea was filled with these creatures, 
with each one making more of its own kind. 
He also made every bird that flies, 
and each bird made more of its own kind. 
The god saw that it was good. 
The god blessed them and said, 
\enquote{Be fruitful, multiply, and fill the water of the seas, 
and let the birds grow in number on the earth.} 
There was an evening, and there was a morning---the fifth day.

Then the god said, 
\enquote{Let the land make living beings according to their kind: 
livestock, creeping things, and wildlife, each according to their kind.}
So it was. 
The god made the wildlife according to their kind, 
the livestock according to their kind, 
and the creatures that creep along the ground according to their kind. 
The god saw that it was good. 

Then the god said, 
\enquote{Let us make humankind in our image, after our likeness, 
so they may rule over the fish of the sea and the birds of the sky, 
over livestock and over all of the earth, 
and over creeping creatures on the earth.} 
The god created humankind in his own image; 
in the image of the god he created them;
male and female he created them. 
The god blessed them, saying, 
\enquote{Be fruitful and multiply! 
Fill the earth and exploit it to your own benefit! 
Rule over the fish of the sea and the birds of the sky 
and every being that moves on the ground.} 

The god said, 
\enquote{Listen. 
I bestow onto you every seed-bearing plant on the surface of the entire earth 
and every tree that has fruit with seed in it. 
They will be yours to eat. 
And to all the animals of the earth, 
to every bird of the sky, 
to all the beings that creep on the ground, 
to all that has the breath of life in it, 
every green plant is your food.} 
So it was. 
The god saw all that he had made and it was \textit{very} good. 
There was an evening, and there was a morning---a sixth day.

And so the sky and the earth were done, with all their units. 
On the seventh day, the god finished the work he had been doing, 
and he stopped all his labour. 
The god blessed the seventh day and separated it 
because on it he rested from all the work he had been doing in creation.

These are the generations of the sky and the earth. 
