\chapter*{Chapter Two}
\addcontentsline{toc}{chapter}{Chapter Two}

When the god Yahweh began to create the earth and the sky, no plants had yet 
sprouted, because the god Yahweh had not made it rain on earth, 
and there was no one to till the ground.  Springs would rise from the earth 
and water the surface of the ground.

Then the god Yahweh took some soil from the ground and formed from it a man.  
He breathed into his nostrils the breath of life, and the man became 
a living being.  The god Yahweh made an orchard in the east, in a place called 
Eden, and he put the man in there.  The god Yahweh made every tree that was beautiful 
and every tree that was edible from the soil.  In the middle of the orchard was 
the tree of life and the tree of the understanding of good and evil.

A river flows from Eden to water the orchard, and from there it branches out 
into four.  The name of the first is Pishon, which runs through the entire land 
of Havilah, where there is gold.  The gold of that land is excellent; in there is 
bdellium and onyx.  The name of the second is Gihon, the same as the one that 
winds through the whole land of Cush.  The name of the third is Tigris, the one 
that flows out to the east of Assyria.  The fourth is Euphrates.

Then the god Yahweh took the man and put him in the orchard in Eden to till it 
and to protect it.  The god Yahweh commanded the man, saying, 
\enquote{Every fruit of every tree in the orchard, you may freely eat, but 
those of the tree of the understanding of good and evil, you must not eat 
because if you eat from it, you will absolutely die.}

Then the god Yahweh said, \enquote{It is not right for the man to be alone.  
I will make a companion who is suitable for him.} And out of the soil the god 
Yahweh made every living animal on land and every bird in the sky.  He brought 
them to the man to see what he would name them.  Whatever name the man gave, 
that became its name.  So the man gave names to all the livestock, to the birds 
in the sky, and to all wildlife, but no suitable companion was found for Adam.

So the god Yahweh caused the man to fall into a deep sleep, and he slept.  He 
took a part of the man's side and closed the place in the flesh.  Then the god 
Yahweh made a woman from the part he had taken out of the man, and he brought 
her to the man.  Then the man said, \enquote{This, finally, is bone of my bones 
and flesh of my flesh; this one will be named “woman”, because she was taken out 
of man.} And it was so that when a man leaves his father and mother, when he 
unites with his wife, they become one flesh.  

The man and his wife were naked, but they were not ashamed.
