\chapter*{Chapter Five}
\addcontentsline{toc}{chapter}{Chapter Five}

This is the record of the generations of Adam.

When the god created humanity,
he made them in his likeness:
male and female, he created them.
And when they were created,
he blessed them 
and called them Human.

When Adam had lived one hundred and thirty years,
he had a son in his likeness,
after his image, 
and called him Seth.
After the birth of Seth,
Adam lived eight hundred years  more
and had sons and daughters.
Adam lived a total of
nine hundred and thirty years,
and then he died.

When Seth had lived one hundred and five years, 
he had Enosh. 
After the birth of Enosh,
Seth lived eight hundred and seven years more 
and had sons and daughters. 
Seth lived a total of 
nine hundred and twelve years,
and then he died.

When Enosh had lived ninety years, 
he had Kenan. 
After the birth of Kenan,
Enosh lived eight hundred and seven years more 
and had sons and daughters. 
Enosh lived a total of 
nine hundred and twelve years,
and then he died.

When Kenan had lived seventy years,
he had Mahalalel.
After the birth of Mahalalel,
Kenan lived eight hundred and forty years more
and had sons and daughters.
Kenan lived
a total of nine hundred and ten years,
and then he died.

When Mahalalel had lived sixty-five years,
he had Yared.
After the birth of Yared,
Mahalalel lived eight hundred and thirty years more
and had sons and daughters.
Mahalalel lived
a total of eight hundred and ninety-five years,
and then he died.

When Yared had lived one hundred and sixty-two years,
he had Enoch.
After the birth of Enoch,
Yared lived eight hundred years more
and had sons and daughters.
Yared lived
a total of nine hundred and sixty-two years,
and then he died.

When Enoch had lived sixty-five years,
he had Methuselah.
After the birth of Methuselah,
Enoch walked with the god for three hundred years,
and he had sons and daughters.
Enoch lived
a total of three hundred and sixty-five years. 
Enoch walked with the god, 
and then he was no more,
for the god took him.

When Methuselah had lived one hundred and eighty-seven years,
he had Lamech.
After the birth of Lamech,
Methuselah lived seven hundred and eighty-two years more
and had sons and daughters.
Methuselah lived
a total of nine hundred and sixty-nine years,
and then he died.

When Lamech had live one hundred and eighty-seven years,
he had Noah.\footnote{
    The original reads,
    `When Lamech had live one hundred and eighty-seven years,
    he had a son,
    and he named him Noah.
    ``This one,'' he said, 
    ``will provide us relief from our work
    and from the toil of our hands,
    out of the very soil which Yahweh has cursed.'''
    There is an opinion that this etymological explanation 
    is inserted by a nonpriestly writer,
    but my omission is largely a stylistic one,
    to keep this chapter's established style.
}
After the birth of Noah, 
Lamech live five hundred and ninety-five years more 
and had sons and daughters. 
Lamech lived
a total of seven hundred and seventy-seven years,
and then he died.

When Noah had live one hundred and eighty-seven years,
he had Shem, Ham, and Japheth.
