\chapter*{Chapter Six}
\addcontentsline{toc}{chapter}{Chapter Six}

When humanity began to populate the earth,
daughters were born to them.
The sons of gods saw how beautiful
the daughters of humanity were 
and took those they wanted as wives.
Yahweh said, `My breath will not stay in humanity forever,
for they are only flesh.
Let their days be one hundred and twenty years.'
It was then and some time after
that the Nephilim appeared on earth,
when these gods came in unto the daughters of humanity,
who gave them children.
The same children became the heroes of old,
the men of renown.

Yahweh saw that the wickedness
of humanity was great
and that every thought created by their mind
was nothing but continuous evil.
Yahweh regretted that he made humanity on earth,
and his heart was saddened.
He said, `I will erase from the earth 
the humans that I created,
humans, together with the animals, insects, and the birds in the sky,
for I regret that I made them.'
But Yahweh looked favourably unto Noah.

\secsep

\noindent This is the story of Noah. 
He was righteous and blameless.
Noah walked with the god. 
He had three sons: Shem, Ham, and Yafet.

The earth had become corrupt in front of the god;
the earth was filled with violence.
When the god saw the corruption of the earth,
for all flesh had violated its assigned roles on earth,
the god said to Noah, `I will bring an end to all flesh,
as the earth has been filled with violence because of them.
I am about to destroy them along with the earth.

`Make yourself an ark of gopher wood.
Make it with compartments, 
and cover it inside and out with pitch.
This is how you make it:
make the length of the ark three hundred cubits,
its width fifty cubits, 
and its height thirty cubits.
Make an opening for daylight, 
and finish it a cubit higher than the top.
Put the entrance on its side.
And make first, second, and third levels.

`As for me,
I am about to bring a flood of the waters on the earth 
to destroy all flesh under the sky that has the breath of life.
Everything will die.
And then I will establish my covenant with you.

`You will enter the ark, 
with your sons,
your wife,
and your sons' wives. 
From every living thing,
from all flesh,
take two of each and bring them into the ark 
to keep alive with you;
they will be male and female---birds of every kind,
livestock of every kind,
every creeping thing on earth,
two of each will come to you to stay alive.

`For your part,
take everything that is eaten and stor it away,
to serve as food for you 
and for them.'

Noah did so, just as the god commanded him.