\chapter*{Chapter One}
\addcontentsline{toc}{chapter}{Chapter One}

The words of Kohelet, son of David, king in Jerusalem:

`Absolute absurdity,' says Kohelet. `Absolute absurdity. Everything is absurd. What good does anyone gain from all their labours under the sun? 

`A generation goes, another comes, but the earth remains forever. The sun rises, and the sun sets, and it hurries back to where it rises. The wind blows southward, turns northward, and goes round and round; round and round the wind goes, and on its rounds it returns. All the streams flow into the sea, but the sea is never full; to the place from which they flow, the streams flow back again.

`All of this is wearisome. No one can describe it. The eye is never satisfied with seeing, nor is the ear ever content with hearing.

`All that is, is all that will be, and all that has been done are what will be done. There is nothing new under the sun. Some might say of something, ``Look, this is new!'' but in reality, that thing has always been here, in the ages before us. There is no remembrance for the past, nor is there for the future. There is no remembrance among those who live in the future.

`I, Kohelet, was king in Jerusalem over Israel. I set my mind to carefully scout and survey all that has been accomplished under the sun. God has given the sons of men an unhappy business that keeps them occupied. I looked at everything done under the sun and saw that it was all absurd, like chasing the wind! What is bent cannot be straightened, and what is lacking cannot be counted.

`I thought to myself, “I have gained magnificence and have gained more wisdom than all who were before me over Jerusalem. My mind has seen much wisdom and knowledge.” So I decided to understand wisdom and knowledge from madness and folly, but even this is like chasing the wind. For with great wisdom comes great frustration.

`Whoever gains more knowledge also gains more heartache.'
