\chapter{Cain and Abel}

Now Adam has had sex with his wife Eve,
and she conceived and gave birth to Cain, saying,
`I have gained a son with the help of Yahweh!'
Later she gave birth to his brother
and named him Abel.
Abel grew up to be the keeper of sheep, 
and Cain grew up to the keeper of the soil.

In due course, 
Cain brought some fruit from the ground 
as a tribute to Yahweh. 
For his part,
Abel brought fat from 
the firstborn of his flock.
Yahweh looked favourably unto Abel and his tribute,
but he did not look favourably unto Cain and his tribute.
So Cain became furious, and he felt abandoned.

Yahweh asked Cain,
`Why are you angry?
Why do you look so unhappy?
If you do things correctly,
will I not accept you?
But if you do them incorrectly,
sin waits at the door.
Sin desires to control you,
but you can be its master.'

Cain talked to his brother Abel.\footnote{
    Some manuscripts of the bible include a quote from Cain.
    I'm taking a liberty to omit it, as I believe, in this case,
    that the Masoretic Text offers a more suspenseful narrative
    than what is offered by 
    the Samaritan, Greek, and Latin manuscripts.
    If I were to include it, I would translate it, 
    `Cain talked to his brother Abel and whispered, 
    ``Come, let us go out into the fields.'''
}

\subsubsection{\secsep}

 While they were in the fields, 
Cain attacked his brother Abel and killed him.

\subsubsection{\secsep}

 Yahweh said to Cain,
`Where is your brother Abel?'

He replied,
`I don't know. 
Am I my brother's keeper?'

`What have you done?!
Listen!
Your brother's blood 
calls out to me from the ground!
Because of this, you are exiled from the ground 
which had opened its mouth
to receive your brother's blood 
from your hands.
If you till the soil,
it will no longer yield to your strength.
You shall be a vagabond on the earth.'

`My punishment is greater than I can bear,'
Cain said to Yahweh.
`You have banished me from the land
and from your presence. 
You have made me a vagabond.
Anyone who finds me will kill me!'

`If anyone kills Cain, vengeance will befall on them sevenfold.'
Yahweh branded Cain, to warn anyone who might try to kill him.

Cain left the presence of Yahweh
and settled in the land of Nod, east of Eden.

\subsubsection{\secsep}

 Now Cain has had sex with his wife,
and she conceived and gave birth to Enoch.
Then Cain founded a city, 
which he named Enoch,
after his son.
Enoch had Irad, 
and Irad had Mehijael,
and Mehijael had Methusael,
and Methusael had Lamech.
Lamech had two wives; 
one was named Adah,
and the other was named Zillah.
Adah gave birth to Jabal,
who was the ancestor of those 
who live in tents and amidst herds,
and his brother Jubal, 
who was the ancestor of 
all who play the lyre and the pipe.
As for Zillah, 
she gave birth to Tubal-cain,
who made copper and iron tools,
and his sister Naamah.

One day, Lamech said to his wives,
`Adah and Zillah, hear my voice.
Oh wives of Lamech, listen to my words.
I have killed a man for wounding me
and a child for bruising me.
If Cain is avenged sevenfold,
than Lamech is avenged seventy-sevenfold.'

\subsubsection{\secsep}

 Adam had sex with his wife again, 
and she gave birth to another son.
She named him Seth because `The god,' she said,
`has sent me another son in place of Abel,
who who killed by Cain.'
Seth had a son who he named Enosh.
It was then that men began to invoke Yahweh by name.