\chapter{The Deluge}

This is the record of the generations of Adam.

When the god created humanity,
he made them in his likeness;
male and female, he created them.
And when they were created,
he blessed them 
and called them human.

When Adam had lived one hundred and thirty years,
he had a son in his likeness,
looking like him, 
and called him Seth.
After the birth of Seth,
Adam lived eight hundred years more
and had sons and daughters.
Adam lived a total of
nine hundred and thirty years,
and then he died.

\generateRecord{Seth}{105}{Enosh}{807}{912}

\generateRecord{Enosh}{90}{Kenan}{815}{905}

\generateRecord{Kenan}{70}{Mahalalel}{840}{910}

\generateRecord{Mahalalel}{65}{Yared}{830}{895}

\generateRecord{Yared}{162}{Enoch}{800}{962}

When Enoch had lived \numspell{65} years,
he fathered Methuselah.
After the birth of Methuselah,
Enoch walked with the god for \numspell{300} years,
and he fathered sons and daughters.
Enoch lived
a total of \numspell{365} years. 
Enoch walked with the god, 
and then he was no more,
for the god took him.

\generateRecord{Methuselah}{187}{Lamech}{782}{969}

\generateRecord{Lamech}{182}{Noah}{595}{777}\footnote{
    The original would read,
    `When Lamech had lived \numspell{182} years,
    he had a son,
    and he named him Noah.
    ``This one,'' he said, 
    ``will provide us relief from our work
    and from the toil of our hands,
    out of the very soil which Yahweh has cursed.''\dots'
    My omission is largely a stylistic one,
    to keep this chapter's established style.
}

When Noah had lived five hundred years,
he fathered Shem, Ham, and Yafet.

\subsubsection*{\secsep}

This is the story of Noah; he walked with the god.\footnote{
    Before and after this are short asides and parentheticals
    which I have found disruptive to the flow
    and have omitted from the text. 
} 

The earth had become corrupt in front of the god;
the earth was filled with violence.
When the god saw the corruption of the earth,
for all flesh had violated its roles on earth,
the god said to Noah, `I will bring an end to all flesh,
as flesh has caused the earth to be filled with violence.
I am about to destroy them along with the earth.

`Make yourself an ark of gopher wood.
Make it with rooms, 
and coat it inside and out with pitch.
The dimensions of the ark should be as follows: 
the length is three hundred cubits, 
the width fifty cubits,
and the height thirty cubits.
Include an opening to allow daylight,
and finish it a cubit higher than the top.
Place the entrance on its side, 
and create first, second, and third levels."

`I am about to bring a flood of the waters upon the earth 
to destroy all flesh under the sky that has the breath of life.
Everything will die.
And then I will establish my pact with you.

`You will enter the ark
with your sons,
your wife,
and your sons' wives. 
From every living thing,
from all flesh,
take two of each and bring them into the ark 
to keep alive with you;
they will be male and female---birds of every kind,
livestock of every kind,
every creeping thing on earth,
two of each will come to you to stay alive.

`Take everything that is to be eaten and store it away
to serve as food for you 
and for them.'

Noah did so, just as the god commanded him.

\section*{\secsep}
On the seventeenth day of the second month
of the six hundredth year of Noah's life,
on that very day, all of the fountains of 
the deep waters
burst apart,
and the windows of the sky broke open.\footnote{
Unfortunately due to the volume of text that I have
omitted and moved to improve the flow of the story, 
I will not be including the omitted verses here in full.
For reference, here are the traditional indeces 
of the omitted verses 7:1--10, 7:12, fragments of 7:16, 7:17, a fragments of 7:19, and 7:23.}
On that very day, Noah, his sons Shem, Ham, and Yafet,
Noah's wife, and wives of his three sons entered the ark,
along with
all the living creatures after its kind,
every kind of land animal after its kind, 
creatures that creep on the earth after its kind,
and every bird after its kind.
Pairs of all flesh, male and female,
that have the breath of life in it came to Noah, to the ark,
just as the god commanded.
Then Yahweh shut him in.

The waters rose and multiplied on the earth,
the ark floated on the surface of the water,
and all the high mountains under the sky were covered.
The waters rose fifteen cubits higher than the mountains. 
All the flesh that moved on the earth died---birds,
wildlife, livestock, all the creatures that swarm above the earth,
and all of humanity.
Everything that had the breath of life in its nostrils, 
everythin on dry land,
died.

The waters reigned over the earth for \numspell{150} days.


\subsubsection{\secsep}

The god kept in mind Noah and all the living beings 
and the animals that were with him in the ark,
and the god caused a wind to move across the earth,
and the waters calmed.
He closed the fountains of the deep waters 
and the windows of the sky.
The waters receded at the end of a hundred and fifty days. 
The ark rested on the mountains of Ararat 
on the seventh day of the seventh month.
The waters continued to recede until the tenth month;
on the first day of the tenth month, 
the mountaintops became visible.

Then Noah opened the window of the ark that he had made
and sent out a dove to see 
whether the waters had dried up from the earth.\footnote{The raven Noah releases is omitted for the flow of the story.}
But the dove could not find a resting place for its foot
and returned to the ark, 
for there was water all over the earth.
Noah waited for seven days and sent out another dove.
Towards the evening,
the dove came back to him with an olive leaf on its bill.
This was how Noah knew that the waters had receded from the earth.
He waited seven days more and sent out the dove,
and it did not return to him anymore. 

On the first day of the first month of the six hundred and first year,
the waters began to dry from the earth,
and when Noah removed the covering of the ark,
he saw that the surface of the earth was drying up. 
On the twenty-seventh day of the secon month,
the earth was dry.

God said to Noah, 
`Go out of the ark, you, your wife, your sons, and your sons' wives.
Bring out every living being of all flesh that is with you:
birds, animals, and everything that creeps on the earth.
Let the swarm on the earth and be fertile and multiply on the earth.'

Noah, along with his sons and his wife and his sons' wives, went
out. Every animal, all the creeping creatures, every bird, everything
that moves on the earth, came out of the ark according to their families.

\footnote{Verses 8:20--22 are omitted.}
And the god blessed Noah and his sons, and he said, 
`Be fertile, multiply, and replenish the earth. 
Every land animal,
every bird of the sky,
all that move on the earth, 
over every fish of the sea 
will have fear and terror for you. 
They are now given into your hands. 
Every moving thing that lives will be yours to eat,
just like the green plants;
I am giving all of this to you.

`You must not, however, eat flesh with its life, its blood.
I will require blood for blood.
I will require the blood, the life, of any animal that kills a person.
I will require the blood, the life, of any person who kills another person.
Whoever sheds the blood of human will have his blood be shed by a human 
because the god made human in his own likeness.
As for you, be fertile, multiply, and swarm over the earth.'

The god said to Noah and to his sons, 
`I am establishing my pact with you and with your descendants, 
and with every living being that is with you---birds, livestock, and every land animal;
those who came out of the ark with you, 
every living being on earth. 
I am establishing my pact with you: 
never again will all
flesh be cut off by the waters of a flood, 
and never again will a flood destroy the earth.

`This is the sign that I set for the pact 
between me and all of you for all ages to come.
I have set my bow in the clouds. 
When I bring clouds over the earth and it appears in the clouds,
I will remember our pact.'\footnote{The verses
pertaining to the curse of Canaan are omitted.}

After the flood, Noah lived for \numspell{350} years more. 
Noah lived for a total of \numspell{950} years,
and then he died.

