\chapter{Creation}

When the god began to make the sky and the earth, 
the earth was a formless nothingness, 
darkness covered the surface of the deep waters,
and the breath of the god brushed over the surface of the waters. 

Then the god said, 
`Let there be light.'
Then there was light. 
The god saw that the light was good. 
The god separated the light from the darkness. 
The god named the light `day' and the darkness he named `night'. 
There was an evening, and there was a morning---one day.

Then the god said, 
`Let there be a cover in the waters that will separate water from water.'
Then the god made the cover and separated the water that was under the cover from the water that was above the cover. 
So it was. 
The god named the cover `sky'. 
There was an evening, and there was a morning---a second day.

Then the god said, 
`Let the water under the sky gather into one place, so that dry land may appear.' 
So it was. 
The god named the dry land `earth' and the gathered waters `sea'. 
The god saw that it was good. 

Then the god said, 
`Let the earth bring forth plants: 
every kind that makes grains for its seeds and 
every kind that makes fruits that have in them seeds.'
So it was. 
Thus the earth brought forth plants: 
every kind that makes grains for its seeds and 
every kind that makes fruits that have in them seeds, 
and the god saw that it was good. 
There was an evening, and there was a morning---a third day.

Then the god said, 
`Let there be lights in the sky, 
to separate day from night, and 
let them be for signs of seasons and days and years, 
and let them be the lights in the sky to illuminate the earth.' 
So it was. 
The god made the two large lights---the bigger one to rule the day 
and the smaller one to rule the night---along with the stars. 
The god put all these in the sky to illuminate the earth, 
to rule over the day and over the night, 
and to separate light from darkness. The god saw that it was good.
There was an evening, and there was a morning---a fourth day.

Then the god said, 
`Let the waters swarm with swarms of living beings, 
and let birds fly in the sky above the earth.' 
The god shaped large sea monsters,
and swarms of every kind of living being in the waters,
and every kind of winged bird. 
The god saw that it was good. 
The god blessed them and said, 
`Be fertile, multiply, and fill the water of the seas, 
and let the birds grow in number on the earth.' 
There was an evening, and there was a morning---a fifth day.

Then the god said, 
`Let the land make every kind of living beings: 
livestock, creeping things, and wildlife of every kind.'
So it was. 
The god shaped every kind of wildlife, 
every kind of livestock, 
and every kind of creeping creatures. 
The god saw that it was good. 

Then the god said, 
`Let us make humanity in our likeness, looking like us, 
and let them rule over the fish of the sea and the birds of the sky, 
over livestock over all of the earth, 
and over creeping creatures on the earth.' 
The god created humanity in his own likeness; 
in the likeness of the god he created them;
male and female he created them. 
The god blessed them, saying, 
`Be fertile and multiply! 
Fill the earth and exploit it to your own benefit.
Rule over the fish of the sea and the birds of the sky 
and every being that moves on the ground.

`Listen. 
I bestow onto you every plant that makes grains for its seed 
on the surface of the entire earth 
and every tree that makes fruit with seed in it. 
They will be yours to eat. 
And to all the animals of the earth, 
to every bird of the sky, 
to all the beings that creep on the ground, 
to all that has the breath of life in it, 
every green plant is your food.'
So it was. 
The god saw all that he had made and it was \textit{very} good. 
There was an evening, and there was a morning---a sixth day.

And so the sky and the earth were done, with all their units. 
On the seventh day, the god finished the work he had been doing, 
and he stopped all his labour. 
The god blessed the seventh day and separated it 
because on it he rested from all the work he had been doing in creation.

These are the generations of the sky and the earth. 
