\chapter{The First Humans}

When the god Yahweh began to create the earth and the sky, no plants had yet 
sprouted, because the god Yahweh had not made it rain on earth, 
and there was no one to till the ground.  Springs would rise from the earth 
and water the surface of the ground.

Then the god Yahweh took some soil from the ground and made from it a man.  
He breathed into his nostrils the breath of life, and the man became 
a living being.  The god Yahweh made an orchard in the east, in a place called 
Eden, and he put the man in there.  The god Yahweh made every tree that was beautiful 
and every tree that was edible from the soil.  In the middle of the orchard was 
the tree of life and the tree of the understanding of good and evil.

A river flows from Eden to water the orchard, and from there it branches out 
into four.  The name of the first is Pishon, which runs through the entire land 
of Havilah, where there is gold.  The gold of that land is excellent; in there is 
bdellium and onyx.  The name of the second is Gihon, the same as the one that 
winds through the whole land of Cush.  The name of the third is Tigris, the one 
that flows out to the east of Assyria.  The fourth is Euphrates.

Then the god Yahweh took the man and put him in the orchard in Eden to till it 
and to protect it.  The god Yahweh commanded the man, saying, 
\enquote{Every fruit of every tree in the orchard, you may freely eat, but 
those of the tree of the understanding of good and evil, you must not eat 
because if you eat from it, you will absolutely die.}

Then the god Yahweh said, \enquote{It is not right for the man to be alone.  
I will make a companion who is suitable for him.} And out of the soil the god 
Yahweh made every living animal on land and every bird in the sky.  He brought 
them to the man to see what he would name them.  Whatever name the man gave, 
that became its name.  So the man gave names to all the livestock, to the birds 
in the sky, and to all wildlife, but no suitable companion was found for Adam.

So the god Yahweh caused the man to fall into a deep sleep, and he slept.  He 
took a part of the man's side and closed the place in the flesh.  Then the god 
Yahweh made a woman from the part he had taken out of the man, and he brought 
her to the man.  Then the man said, \enquote{This, finally, is bone of my bones 
and flesh of my flesh; this one will be named “woman”, because she was taken out 
of man.} And it was so that when a man leaves his father and mother, when he 
unites with his wife, they become one flesh.  

The man and his wife were naked, but they were not ashamed.

\subsubsection*{\secsep}

Now, the snake was more cunning than any wildlife that the god Yahweh made. 
It said to the woman, 
\enquote{Is it really true that the god said, 
    \enquote{You must not eat from any tree in the orchard}?}

And the woman said to the snake, 
\enquote{We may eat them, but as to the fruit of the tree in the middle of the orchard, 
the god said, \enquote{You must not eat from it, and you must not touch it, 
or else you will die.}}

\enquote{Surely you will not die because the god knows that when you eat from it, 
your eyes will open and you will be like gods, 
who understand good and evil.}

And when the woman saw that the tree produced fruit that 
was edible, attractive, and able to make one have that understanding, 
she took some of its fruit and ate it. 
She also gave some to her husband, and he ate it. 
Their eyes opened, and they realised that they were naked. 
So they sewed fig leaves together and made coverings for themselves.


\subsubsection*{\secsep} 
They heard the sound of the god Yahweh moving in the orchard 
in the breezy part of the day, 
and they hid from the god Yahweh among the trees of the orchard. 
But the god Yahweh called to the man and said, \enquote{Where are you?}

The man replied, 
\enquote{I heard your sound in the orchard, 
and I was afraid because I was naked, 
and so I hid.}

\enquote{Who told you that you were naked? 
Did you eat from the tree which I commanded you not to eat?}

\enquote{That woman you gave me---she gave me some, and I ate it.}

So the god Yahweh said to the woman, \enquote{What in the world have you done?!}

And the woman replied, \enquote{The snake lied to me, and I ate.}

The god Yahweh said to the snake, 
\enquote{Because you have done this, 
you are cursed above all wildlife and livestock.
You will crawl on your belly, 
and you will eat soil 
for all the days of your life. 
I will put hostilities between you and the woman 
and between your children and her children. 
Her children will attack your head, 
and you will attack her children's heel.} 

To the woman, he said, 
\enquote{I will intensify the pains of your childbearing. 
With suffering you will give birth to children. 
You will be longing for your husband, but he will dominate you.} 

To Adam, he said, 
\enquote{Because you would rather listen to the voice of your wife 
and eat from the tree which I have forbidden, 
cursed is the ground because of you. 
With suffering shall you eat from it all the days of your life. 
It will produce thorns and thistles for you, 
and you will eat the plants of the field. 
By the sweat of your brow shall you eat your food 
until you return to the soil, for from it you were taken. 
You are dust, and to dust you will return.}

The man named his wife Eve, because she was the mother of all the living. 
The god Yahweh made clothes from animal skins for Adam and his wife 
and then he clothed them.

And the god Yahweh said, 
\enquote{Listen. 
The man has become like one of us, 
understanding of good and evil. 
He must not be allowed to reach his hand out 
and take also from the tree that gives life and eat and live forever.}

So the god Yahweh sent him from the orchards of Eden 
to till the ground from which he had been taken.

He drove him out and put at the east of Eden
the Cherubim and a flaming sword which was ever-turning, 
to guard over the way to the tree of life.

